---
title: "DATA 607: Assignment 3"
author: "Farhana Akther"
output: html_document
date: "2023-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Question 1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

### Load Libraries

```{r}
library(tidyverse)
```

#### Load data from Github
```{r}
College_maj<-read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')
```

#### Code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r}
find='DATA|STATISTICS'
College_maj_sub <- College_maj$Major[grep(find, College_maj$Major)]
print(College_maj_sub)
```

grep() function has been used to search for matches of a pattern within each element of the given string.

### Question 2

#### Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

#### Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r}
Fruits <- data.frame(c("bell pepper", "bilberry", "blackberry","blood orange","blueberry","cantalope","chili pepper","cloudberry","elderberry","lime","lychee","mulberry","olive","salal berry"))
cat(paste0(Fruits), collapse=",")
```


### Question 3

#### Describe, in words, what these expressions will match:

1. (.)\1\1 

2. "(.)(.)\\2\\1"

3. (..)\1

4. "(.).\\1.\\1"

5. "(.)(.)(.).*\\3\\2\\1"

#### Answers: 


1.
(.)\1\1:  Regular expression is used to match a pattern in a strings with character repeats three times or more in it.

```{r}
exp <- c("toooo little", "sooo cute", "blackberry", "blackberrrry", "limeee", "lime", "12345", "347565", "07770")
str_subset(exp, "(.)\\1\\1")
```
2.
"(.)(.)\\2\\1": Regular expression is used to match strings with a set of 4 characters with 2 characters attached to the same 2 characters in reverse order(e.g: bell pepper and chili pepper).

```{r}
str_view(fruit, "(.)(.)\\2\\1")
```

3.
(..)\1: This regular expression is used to match any strings that have a repeated pair of letters (e.g.: banana and/or coconut). 

```{r}
str_view(fruit, "(..)\\1")
```

4.
"(.).\\1.\\1":  Regular expression is used to match any strings that has the same character repeat 3 times and they are all separated by one character (e.g: banana and papaya).

```{r}
str_view(fruit, "(.).\\1.\\1")
```

5.
"(.)(.)(.).*\\3\\2\\1": This regular expression is used to match any strings with any 3 characters that repeat in the reverse order (e.g:347743 and/or abcdeffedcba)

```{r}
exp <- c("toooo little", "abcdeffedcba", "blackberry", "blackberrrry", "limeee", "lime", "12345", "07770", "077770", "347743", "34788743")
str_subset(exp, "(.)(.)(.).*\\3\\2\\1")
```


### Question 4

#### Construct regular expressions to match words that:

1. Start and end with the same character.

2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

3. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

Answers:

1. Start and end with same character:
```{r}
df.names <-c("alisha", "farhana", "anna", "sahana", "church", "bob", "harry", "paul", "eleven", "bubble", "cell", "apple", "dog", "ada", "sense", "banana", "pepperoni", "india", "ten", "twelve", "soso", "oso", "bandana", "Louisiana", "Missouri", "Mississippi", "Connecticut", "google", "conscience", "dalda", "short", "Evon", "ele", "Tort")

regex_expr1 <-"^(.)((.*\\1$)|\\1?$)"
str_subset(df.names,regex_expr1)
```

2. Contain a repeated pair of letters:
```{r}
regex_expr2 <-"([A-Za-z][A-Za-z]).*\\1"
str_subset(df.names,regex_expr2)
```

3. Contain one letter repeated in at least three places
```{r}
regex_expr3 <-"([A-Za-z]).*\\1.*\\1"
str_subset(df.names,regex_expr3)
```


