---
title: "DATA607: Assignment 6"
author: "Farhana Akther"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web APIs

### Overview

In this assignment we are given The New York Times  **[APIs]( https://developer.nytimes.com/api)** web page to sign up for an API key in order to get access to their APIs Data. We will choose one of the APIs, construct an interface in R to read in the JSON data, and transform it into an R Dataframe.

#### Load Libraries:

```{r}
library(dplyr)
library(jsonlite)
library(DT)
```

#### Obtaining API from NYT Developers Web Page
*here we are extracting the data from NYT Book API by calling the Example Calls (2nd URL)*  

```{r}
url <- "https://api.nytimes.com/svc/books/v3/reviews.json?author=Stephen+King&api-key=SOfRrxpoYgeFDX8eoGRqvq6lMEcqyEWH"
```

#### Reading the JSON File     

```{r}
books_nyt <- fromJSON(url, flatten = TRUE)
```

## Create Dataframe

```{r}
books_nyt2 = data.frame(books_nyt$results)
head(books_nyt2)
```

## Transform Data

```{r}
books_nyt3 <- select(books_nyt2, url, publication_dt, byline, book_title, book_author, summary)
colnames(books_nyt3) = c("Article URL", "Publication Date", "Byline", "Book Title", "Book Author", "Summary")
datatable(books_nyt3)
```

## Rearrange Columns

```{r}
books_nyt3 <- books_nyt3[c("Book Title", "Book Author", "Publication Date", "Article URL", "Summary")]
datatable(books_nyt3)
```


## Conclusion

In this assignment, we are able to learn:

* How to create an API account on the New York Times Developer web site.

* How to construct an interface in R to read the JSON data and transform it into a R DataFrame by using fromJSON() to Get a JSON file from NYT Archives API.

* How to create a datafram using data.frame(), transform and rearrange the table by using select() and colnames() functions. 



